A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-02-20 main.tex
Abstract and Intro
3 keywords

This paper presents the first full mechanical formalization of the metatheory for
higher-ranked polymorphic type inference. The system that we formalize
is the bidirectional type system by Dunfield and
Krishnaswami (DK). The DK type system has two variants (a declarative
and an algorithmic one) that have been \emph{manually} proven
\emph{sound}, \emph{complete} and \emph{decidable}. We present 
a mechanical formalization in the Abella theorem prover 
of DK's declarative type system with a novel algorithmic system. We have
several reasons to use a new algorithm. Firstly, our new algorithm
employs \emph{worklist judgements}, which allow precise capture of the
scope of variables and simplify the formalization of scoping in a
theorem prover. Secondly, while DK's original formalization comes with very
well-written manual proofs, there are several details missing,
which complicate the task of writing a formal proof. 
Moreover, some proofs/lemmas are wrong and it is
not clear how to replace these by other valid ones. 
\bruno{others?} Despite the use of a different algorithm we prove the
same results as DK, although with significantly different proofs and
proof techniques. Since such type inference algorithms are quite
subtle and have a complex metatheory, we believe that mechanical
formalizations are an important advance in type-inference research.
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords are mandatory in final camera-ready submission

>>>>

This paper presents the first full mechanical formalization of the metatheory for
higher-ranked polymorphic type inference. The system that we formalize
is the bidirectional type system by Dunfield and
Krishnaswami (DK). The DK type system has two variants (a declarative
and an algorithmic one) that have been \emph{manually} proven
\emph{sound}, \emph{complete} and \emph{decidable}. We present 
a mechanical formalization in the Abella theorem prover 
of DK's declarative type system with a novel algorithmic system. We have
several reasons to use a new algorithm. Firstly, our new algorithm
employs \emph{worklist judgements}, which allow precise capture of the
scope of variables and simplify the formalization of scoping in a
theorem prover. Secondly, while DK's original formalization comes with very
well-written manual proofs, there are several details missing,
which complicate the task of writing a formal proof. 
Moreover, some proofs/lemmas are wrong and it is
not clear how to replace these by other valid ones. 
Despite the use of a different algorithm we prove the
same results as DK, although with significantly different proofs and
proof techniques. Since such type inference algorithms are quite
subtle and have a complex metatheory, we believe that mechanical
formalizations are an important advance in type-inference research.
\end{abstract}

@@

%% Keywords
%% comma separated list
\keywords{type inference, higher-rank polymorphism, mechanisation}  %% \keywords are mandatory in final camera-ready submission
